---
title: "2020 West Coast Wildfires"
author: "R. Chris Berg"
date: "9/10/2020"
output: 
  html_document:
    theme: yeti
---

```{r setup, include=FALSE}

library(geojsonio)
library(rnaturalearth)
library(tidyverse)
library(sf)
library(leaflet)

west_coast <-
  ne_states( country = "united states of america" , 
             returnclass = "sf" ) %>%
  filter( name %in% c("California","Oregon","Washington") )

fire_areas <- 
  geojson_read(
    "https://opendata.arcgis.com/datasets/5da472c6d27b4b67970acc7b5044c862_0.geojson", 
    what = "sp"
  ) %>%
  st_as_sf()

```

This is meant to be a resource for looking at the location and extent of the severe wildfires on the United States West coast during the summer of 2020. The map is interactive, using [Leaflet](https://leafletjs.com/) and data from the [National Interagency Fire Center](https://data-nifc.opendata.arcgis.com/datasets/wildfire-perimeters) made available on ArcGIS, and covers the whole United States.

## Interactive Map

Click-and-drag to move around the map, and use the mouse wheel/buttons to zoom. Click on one of the fire shapes to see how many acres it's estimated to cover.

Updated as of `r Sys.time()`

```{r map, echo=F, warning=F, message=F}
fire_areas %>%
  leaflet(width = "100%") %>%
  addProviderTiles(provider = "CartoDB.Positron") %>%
  addPolygons(
    popup = ~paste0(IncidentName, "<br>", "Area (acres): ", prettyNum(GISAcres, big.mark=",")),
    stroke = FALSE,
    smoothFactor = 0,
    fillOpacity = 0.7,
    color = "#F8766D"
  ) %>% 
  setView(lat=44, lng=-123, zoom=7 )
```

